Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-23 article-models-ttinternal.tex
initiated CwF section

\begin{defn}
An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quadruple 
\begin{equation*}
\unfold{\extalg{A}}
\end{equation*}
consisting of a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
a family $\cftalgf{\cftalg{A}}$ over the context
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
a morphism $\cftctxext{\cftalg{A}}$ from $\ctxext{\cftalgc{\cftalg{A}}}{
\cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
$\Gamma$ and a morphism $\cftfamext{\cftalg{A}}$ from 
$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}$ 
to $\cftalgf{\cftalg{A}}$ in context
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, satisfying the judgmental equalities of
\autoref{eq:extalg-eq1,eq:extalg-eq2}.
\end{defn}

@@

\begin{lem}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
Then $\cftctxext{\extalg{A}}$ and $\cftfamext{\extalg{A}}$ 
correctly interpret the rules for context
extension and family extension.

More precisely, for $x_0:\cftalgc{\cftalg{A}}$, $x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
$x_2:\subst{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftalgf{\cftalg{A}}}$ we have the judgmental equality
\begin{align*}
\subst{{x_2}{{x_1}{\cftfamext{\extalg{A}}}}}{{x_0}{\cftctxext{\extalg{A}}}}
& \jdeq
\subst{x_2}{{{x_1}{{x_0}{\cftctxext{\extalg{A}}}}}{\cftctxext{\extalg{A}}}}.
\intertext{and for $y_0:\cftalgf{\extalg{A}}$, $y_1:\subst{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}$ and 
$y_2:\subst{y_1}{{y_0}{\jcomp{}{\cftctxext{\extalg{A}}}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftalgf{\extalg{A}}}}}}$
we have the judgmental equality}
\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext{\extalg{A}}}{\cftfamext{\extalg{A}}}}}}{{y_0}{\cftfamext{\extalg{A}}}}
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\cftfamext{\extalg{A}}}}}{\cftfamext{\extalg{A}}}}.
\end{align*}
\end{lem}

\begin{proof}
Both proofs are simple calculations. For the first judgmental equality we have
\begin{align*}
\subst{{x_2}{{x_1}{\epsilon_1}}}{{x_0}{\epsilon_0}}
& \jdeq
  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\epsilon_1}}}}{\epsilon_0}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{A}}{\epsilon_1}}}}}{\epsilon_0}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{A}}{\epsilon_1}}{\epsilon_0}}}}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}{\epsilon_0}}}}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\epsilon_0}{\idtm{\jcomp{}{\epsilon_0}{P}}}}}}}{\epsilon_0}
  \\
& \jdeq
  \subst{\tmext{\subst{x_1}{{x_0}{\epsilon_0}}}{x_2}}{\epsilon_0}
  \\
& \jdeq
  \subst{x_2}{{{x_1}{{x_0}{\epsilon_0}}}{\epsilon_0}}.
\end{align*}
and to prove the second judgmental equality we calculate
\begin{align*}
\subst{{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}{{y_0}{\epsilon_1}}
& \jdeq
  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}}{\epsilon_1}}
        }
      }
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}{\epsilon_1}}
        }
      }
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\epsilon_1}{\idtm{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}}}}}}{\epsilon_1}
  \\
& \jdeq
  \subst{\tmext{\subst{y_1}{{y_0}{\epsilon_1}}}{y_2}}{\epsilon_1}
  \\
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\epsilon_1}}}{\epsilon_1}}.\qedhere
\end{align*}
\end{proof}

@@

\begin{thm}
Suppose that $(A,P,\epsilon_0,\epsilon_1)$ is an extension algebra in context
$\Gamma$. Then 
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
is an extension algebra in context $\ctxext{\Gamma}{A}$.
\end{thm}

\begin{proof}
We first need to verify that $\jcomp{}{\epsilon_0}{\epsilon_1}$ is a morphism
from $\ctxext{\jcomp{}{\epsilon_0}{P}}{\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}}$
to $\jcomp{}{\epsilon_0}{P}$. This follows from the judgmental equality
$\jcomp{}{\epsilon_1}{{}{\epsilon_0}{P}}\jdeq
\jcomp{}{\epsilon_0}{{}{\epsilon_0}{P}}$, which we have proved in
\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
of exactly the right sort, so the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$
satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxext
  { \jcomp{}{\epsilon_0}{P}
    }
  { { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    { \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{P}
          }
        }
      }
    } 
  \ar{r}
    { \jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
        \jcomp{}{\epsilon_1}{%
          \jcomp{}{\epsilon_0}{\epsilon_1}}}}
  \ar{d}[swap]{
    \jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \idtm{
        \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
          \jcomp{}{\epsilon_1}{%
            \jcomp{}{\epsilon_0}{P}
            }
          }
        }
      }
    }
& \ctxext
    {\jcomp{}{\epsilon_0}{P}}
    {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{d}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \\
\ctxext
  {\jcomp{}{\epsilon_0}{P}}
  {\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{P}}} 
  \ar{r}[swap]{\jcomp{}{\epsilon_0}{\epsilon_1}} 
& \jcomp{}{\epsilon_0}{P}
\end{tikzcd}
\end{equation*}
\end{small}
commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
for the quadruple
$(P,\jcomp{}{\epsilon_0}{P},\epsilon_1,\jcomp{}{\epsilon_0}{\epsilon_1})$. Note
that this follows from \autoref{eq:extalg-eq2} provided that we can show that
\begin{align}
\jvcomp{}{\idtm{\jcomp{}{\epsilon_0}{P}}}{%
  \jcomp{}{\epsilon_1}{%
    \jcomp{}{\epsilon_0}{\epsilon_1}}}
& \jdeq
  \jcomp{}{\epsilon_0}{%
    \jvcomp{}{\idtm{P}}{\jcomp{}{\epsilon_0}{\epsilon_1}}
    }
  \label{eq:extalg-infty1}
  \\
\jvcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
  \idtm{
    \jcomp{}{\jcomp{}{\epsilon_0}{\epsilon_1}}{%
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
& \jdeq
\jcomp{}{\epsilon_0}{%
  \jvcomp{}{\epsilon_1}{%
    \idtm{
      \jcomp{}{\epsilon_1}{%
        \jcomp{}{\epsilon_0}{P}
        }
      }
    }
  }
  \label{eq:extalg-infty2}
\end{align}
Note that \autoref{eq:extalg-infty1} follows if we can show that
\begin{equation*}
\jcomp{}{\epsilon_1}{\jcomp{}{\epsilon_0}{\epsilon_1}}
  \jdeq
  \jcomp{}{\epsilon_0}{\jcomp{}{\epsilon_0}{\epsilon_1}}.
\end{equation*}
This is a special case of \autoref{lem:extalg-twins}. The second judgmental
equality, \autoref{eq:extalg-infty2}, is trivial.
\end{proof}

@@

A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
\begin{equation*}
(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}},\cftempc{\cftalg{A}},
\cftempf{\cftalg{A}})
\end{equation*}
for which 
the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty algebra in context
$\Gamma$, satisfying additional judgmental equalities expressing the 
compatibility of $\cftctxext{\cftalg{A}}$ and $\cftfamext{\cftalg{A}}$ with 
$\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
There will be five such judgmental equalities. Four of them will express that
extension acts as the identity term when applied to the empty object in one of
its arguments. The other expresses that extension of families restricted to
families above the empty object is the same as extension of contexts.

We can immediately state the first two:
\begin{align}
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{extempalg-eq1}
  \\
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext{\cftalg{A}}}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{extempalg-eq2}
\end{align}
To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
we use the following lemma.

@@

\begin{defn}
Let $\cftalg{A}$ and $\cftalg{B}$ be term-algebras in context $\Gamma$. 
A term-operation $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$ is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
consisting of
\begin{align*}
\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
such that the pair $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is both an 
extension homomorphism from
$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext{\cftalg{A}},\cftfamext{\cftalg{A}})$ 
to 
$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftctxext{\cftalg{B}},\cftfamext{\cftalg{B}})$ 
in context $\Gamma$
and an empty homomorphism from 
$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$
to
$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftempc{\cftalg{B}},\cftempf{\cftalg{B}})$
in context $\Gamma$.
\end{defn}

>>>>

\begin{defn}
An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quadruple 
\begin{equation*}
\unfold{\extalg{A}}
\end{equation*}
consisting of a family $\cftalgc{\cftalg{A}}$ over context $\Gamma$, 
a family $\cftalgf{\cftalg{A}}$ over the context
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$,
a morphism $\cftctxext[\cftalg{A}]$ from $\ctxext{\cftalgc{\cftalg{A}}}{
\cftalgf{\cftalg{A}}}$ to $\cftalgc{\cftalg{A}}$ in context
$\Gamma$ and a morphism $\cftfamext[\cftalg{A}]$ from 
$\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}$ 
to $\cftalgf{\cftalg{A}}$ in context
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$, satisfying the judgmental equalities of
\autoref{eq:extalg-eq1,eq:extalg-eq2}.
\end{defn}

@@

\begin{lem}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$.
Then $\cftctxext[\extalg{A}]$ and $\cftfamext[\extalg{A}]$ 
correctly interpret the rules for context
extension and family extension.

More precisely, for $x_0:\cftalgc{\cftalg{A}}$, 
$x_1:\subst{x_0}{\cftalgf{\cftalg{A}}}$ and 
$x_2:\subst{{x_1}{{x_0}{\cftctxext[\extalg{A}]}}}{\cftalgf{\cftalg{A}}}$ 
we have the judgmental equality
\begin{align*}
\subst{{x_2}{{x_1}{\cftfamext[\extalg{A}]}}}{{x_0}{\cftfamext[\extalg{A}]}}
& \jdeq
  \subst
    {x_2}
    {{{x_1}{{x_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
  \intertext{%
and for $y_0:\cftalgf{\extalg{A}}$, 
$y_1:\subst{y_0}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}$ and 
$y_2:
  \subst
    {y_1}
    { {y_0}
      { \jcomp
          {}
          {\cftfamext[\extalg{A}]}
          {\jcomp{}{\cftfamext[\extalg{A}]}{\cftalgf{\extalg{A}}}}
        }
      }
$
we have the judgmental equality}
\subst
  {{y_2}{{y_1}{\jcomp{}{\cftfamext[\extalg{A}]}{\cftfamext[\extalg{A}]}}}}
  {{y_0}{\cftfamext[\extalg{A}]}}
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\extalg{A}]}}}{\cftfamext[\extalg{A}]}}.
\end{align*}
\end{lem}

\begin{proof}
Both proofs are simple calculations. For the first judgmental equality we have
\begin{align*}
\subst{{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}{{x_0}{\cftctxext[\cftalg{A}]}}
& \jdeq
  \subst{\tmext{x_0}{\subst{x_2}{{x_1}{\cftfamext[\cftalg{A}]}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq
  \subst{x_2}{{x_1}{{x_0}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq 
  \subst{{x_2}{{x_1}{{x_0}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{\tmext{\subst{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{x_2}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{x_2}{{{x_1}{{x_0}{\cftctxext[\cftalg{A}]}}}{\cftctxext[\cftalg{A}]}}.
\end{align*}
and to prove the second judgmental equality we calculate
\begin{align*}
\subst{{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}{{y_0}{\cftfamext[\cftalg{A}]}}
& \jdeq
  \subst{\tmext{y_0}{\subst{y_2}{{y_1}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\idtm{P}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}
        }
      }
  \\
& \jdeq
  \subst
    {y_2}
    { {y_1}
      { {y_0}
        {\jcomp{}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}{\cftfamext[\cftalg{A}]}}
        }
      }
  \\
& \jdeq
  \subst{{y_2}{{y_1}{{y_0}{\jvcomp{}{\cftfamext[\cftalg{A}]}{\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{\tmext{\subst{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{y_2}}{\cftfamext[\cftalg{A}]}
  \\
& \jdeq
  \subst{y_2}{{{y_1}{{y_0}{\cftfamext[\cftalg{A}]}}}{\cftfamext[\cftalg{A}]}}.\qedhere
\end{align*}
\end{proof}

@@

\begin{thm}
Suppose that $\extalg{A}$ is an extension algebra in context
$\Gamma$. Then 
\begin{equation*}
\cftalg{F_A}\defeq(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{proof}
In this proof, we shall use the short-hand notations $\cftctxext$ and $\cftfamext$
only to refer to $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$, respectively,
and not to $\cftctxext[\cftalg{F_A}]$ or $\cftfamext[\cftalg{F_A}]$.

We first need to verify that the domain of the morphism 
$\jcomp{}{\cftctxext}{\cftfamext}$ is indeed
$\ctxext{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}$. 
This follows from the judgmental equality
$\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}\jdeq
\jcomp{}{\cftctxext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$, which we have proved in
\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
of exactly the right sort, so the quadruple
$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$
satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxext
  { \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
    }
  { { \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    { \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftfamext}{%
          \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
          }
        }
      }
    } 
  \ar{r}
    { \jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
        \jcomp{}{\cftfamext}{%
          \jcomp{}{\cftctxext}{\cftfamext}}}}
  \ar{d}[swap]{
    \jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
      \idtm{
        \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
          \jcomp{}{\cftfamext}{%
            \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
            }
          }
        }
      }
    }
& \ctxext
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
  \ar{d}{\jcomp{}{\cftctxext}{\cftfamext}}
  \\
\ctxext
  {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
  \ar{r}[swap]{\jcomp{}{\cftctxext}{\cftfamext}} 
& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
\end{tikzcd}
\end{equation*}
\end{small}%
commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
for the quadruple
$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$. Note
that this follows from \autoref{eq:extalg-eq2} provided that we can show that
\begin{align}
\jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
  \jcomp{}{\cftfamext}{%
    \jcomp{}{\cftctxext}{\cftfamext}}}
& \jdeq
  \jcomp{}{\cftctxext}{%
    \jvcomp{}{\idtm{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
    }
  \label{eq:extalg-infty1}
  \\
\jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
  \idtm{
    \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
      \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    }
  }
& \jdeq
\jcomp{}{\cftctxext}{%
  \jvcomp{}{\cftfamext}{%
    \idtm{
      \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    }
  }
  \label{eq:extalg-infty2}
\end{align}
Note that \autoref{eq:extalg-infty1} follows if we can show that
\begin{equation*}
\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftfamext}}
  \jdeq
  \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftfamext}}.
\end{equation*}
This is a special case of \autoref{lem:extalg-twins}. The second judgmental
equality, \autoref{eq:extalg-infty2}, is trivial.
\end{proof}

@@

A term-algebra $\cftalg{A}$ in context $\Gamma$ is going to be a septuple
\begin{equation*}
(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftalgt{\cftalg{A}},
\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}],\cftempc{\cftalg{A}},
\cftempf{\cftalg{A}})
\end{equation*}
for which 
the quadruple $\unfold{\extalg{A}}$ is an extension algebra in context 
$\Gamma$, the quintuple $\unfold{\empalg{A}}$ is an empty algebra in context
$\Gamma$, satisfying additional judgmental equalities expressing the 
compatibility of $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$ with 
$\cftempc{\cftalg{A}}$ and $\cftempf{\cftalg{A}}$.
There will be five such judgmental equalities. Four of them will express that
extension acts as the identity term when applied to the empty object in one of
its arguments. The other expresses that extension of families restricted to
families above the empty object is the same as extension of contexts.

We can immediately state the first two:
\begin{align}
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{extempalg-eq1}
  \\
\jtermeq*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}{\idtm{\cftalgc{\cftalg{A}}}}
  \label{extempalg-eq2}
\end{align}
To see what $\subst{\phi_0}{\epsilon_1}$ can be, we must know its type first.
It is a morphism from $\subst{\phi_0}{\ctxext{P}{\jcomp{}{\epsilon_0}{P}}}$ to
$\subst{\phi_0}{P}$. We already know that $\subst{\phi_0}{P}\jdeq A$ by
\autoref{empalg-eq1} and to compute $\subst{\phi_0}{\jcomp{}{\epsilon_0}{P}}$
we use the following lemma.

@@

\begin{defn}
Let $\cftalg{A}$ and $\cftalg{B}$ be term-algebras in context $\Gamma$. 
A term-operation $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$ is a triple $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}},\cfthomt{\cfthom{f}})$
consisting of
\begin{align*}
\jhom*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{B}}}{\cfthomc{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{\cftalgf{\cftalg{A}}}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
such that the pair $(\cfthomc{\cfthom{f}},\cfthomf{\cfthom{f}})$ is both an 
extension homomorphism from
$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftctxext[\cftalg{A}],\cftfamext[\cftalg{A}])$ 
to 
$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftfamext[\cftalg{B}],\cftfamext[\cftalg{B}])$ 
in context $\Gamma$
and an empty homomorphism from 
$(\cftalgc{\cftalg{A}},\cftalgf{\cftalg{A}},\cftempc{\cftalg{A}},\cftempf{\cftalg{A}})$
to
$(\cftalgc{\cftalg{B}},\cftalgf{\cftalg{B}},\cftempc{\cftalg{B}},\cftempf{\cftalg{B}})$
in context $\Gamma$.
\end{defn}
